---
engine: knitr
---

# Errors and updates {.unnumbered}

::: {.callout-note}
Chapman and Hall/CRC published this book in July 2023. You can purchase that [here](https://www.routledge.com/Telling-Stories-with-Data-With-Applications-in-R/Alexander/p/book/9781032134772).

This online version has some updates to what was printed. An online version that matches the print version is available [here](https://rohanalexander.github.io/telling_stories-published/).
:::

*Last updated: 9 November 2024.*

The book was reviewed by Piotr Fryzlewicz in *The American Statistician* [@Fryzlewicz2024] and Nick Cox on [Amazon](https://www.amazon.com/gp/customer-reviews/R3S602G9RUDOF/ref=cm_cr_dp_d_rvw_ttl?ie=UTF8&ASIN=1032134771). I am grateful that they gave such a lot of their time to provide the review, as well as their corrections and suggestions.

Since the publication of this book in July 2023, there have been a variety of changes in the world. The rise of generative AI has changed the way that people code, Python has become easier to integrate alongside R because of Quarto, and packages continue to update (not to mention a new cohort of students has started going through the book). One advantage of having an online version is that I can make improvements.

I am grateful for the corrections and suggestions of: 
Andrew Black, 
Clay Ford, 
Crystal Lewis, 
David Jankoski, 
Donna Mulkern, 
Emi Tanaka, 
Emily Su, 
Inessa De Angelis, 
James Wade, 
Julia Kim, 
Krishiv Jain, 
Seamus Ross, 
Tino Kanngiesser, and 
Zak Varty.

## Class usage

I have found various ways to use this book in classes. While traditional chalk-and-talk lectures work, if the students can commit to reading the chapter before the class, then I have found that using class for group-based projects and discussion is more enjoyable. Each week create small groups, each of two to four students (randomly create new groups every week to give students a chance to work with new people). Then generally following a "think-pair-share" exercise [@lyman1981responsive] have them work through most exercises, first by themselves, then compare with their group, and finally share selected answers with the class. I recommend creating a Google Doc and using that in places to make it easier to share. If you take this approach then the weekly quiz becomes especially important to ensure students are doing the readings.

In terms of timing and coverage, I have found that if Part I "Foundations" is covered, then the rest of the chapters are fairly independent. While I try to go with one chapter per week, students sometimes take a while to get started, and the first three chapters takes about three weeks (even though there is not much to the first chapter).

Typically, somewhere between the first and second papers is where it all starts to come together. It is important that Paper 1 is returned quickly to them so that they can incorporate lessons from that for future papers.


## Errors

The following errors exist in the print version, but have been updated in the online version. If you notice an error not mentioned below, please submit an [issue](https://github.com/RohanAlexander/telling_stories/issues) or send an email: rohan.alexander@utoronto.ca.


- p.\ xxi: Add Alex Hayes to the Acknowledgments.
- p. 20: Add "packages" to "use the `tidyverse` and `janitor` packages."
- p. 34: `"daily-shelter-overnight-service-occupancy-capacity-2021"` should be `"daily-shelter-overnight-service-occupancy-capacity-2021.csv"` (note the ".csv" added).
- p. 34: Replace the first code chunk with the second:

```{r}
#| eval: false
#| echo: true

toronto_shelters_clean <-
     clean_names(toronto_shelters) |>
     select(occupancy_date, id, occupied_beds)

head(toronto_shelters_clean)
```

```{r}
#| eval: false
#| echo: true

toronto_shelters_clean <-
  clean_names(toronto_shelters) |>
  mutate(occupancy_date = ymd(occupancy_date)) |> 
  select(occupancy_date, occupied_beds)

head(toronto_shelters_clean)
```

- p. 38: "At this point we can make a nice graph of the number of ridings won by each party in the 2019 Canadian Federal Election." should refer to the 2021 election.
- p. 41: Remove a stray ":::".
- p. 66: "New Project$dots" should be "New Project...".
- p. 138: `scale_color_brewer(palette = "Set1")` is unnecessary and should be removed.
- p. 138: The figure caption should refer to inflation not unemployment.
- p. 154: Q9 is missing a "work" after the code chunk and before the "if".
- p. 188: "Leonhard Euler" should be "Carl Friedrich Gauss".
- p. 279: "detonated" should be "denoted".
- p. 342: Q5 option b is repeated in option c.
- p. 347: The "Exploratory data analysis" chapter of *R for Data Science* is 11, not 12.
- p. 353: Fix a "the the".
- p. 355: "...results in an estimate of 5,814, both of which are too low." should be "...results in an estimate of 11,197, the former is too low and the latter is too high."
- p. 371: The sentence that references Figure 11.11a was confusing and needs to more clearly reference the figure. 
- p. 587: The link should be: https://fivethirtyeight.com/features/police-misconduct-costs-cities-millions-every-year-but-thats-where-the-accountability-ends/




