---
engine: knitr
---

# Errata and updates {#sec-errata}

I am grateful for the corrections and suggestions of: 
Andrew Black, 
Clay Ford, 
Crystal Lewis, 
David Jankoski, 
Donna Mulkern, 
Inessa De Angelis, 
James Wade, 
Tino Kanngiesser, and 
Zak Varty.

## Errata

The following errata exist in the print version, but have been updated in the online version. If you notice an error not mentioned below, please submit an [issue](https://github.com/RohanAlexander/telling_stories/issues) or send an email: rohan.alexander@utoronto.ca.

- p. xxi: Add Alex Hayes to the Acknowledgments.
- p. 20: Add "packages" to "use the `tidyverse` and `janitor` packages."
- p. 34: `"daily-shelter-overnight-service-occupancy-capacity-2021"` should be `"daily-shelter-overnight-service-occupancy-capacity-2021.csv"` (note the ".csv" added).
- p. 34: Replace the first code chunk with the second:

```{r}
#| eval: false
#| echo: true

toronto_shelters_clean <-
     clean_names(toronto_shelters) |>
     select(occupancy_date, id, occupied_beds)

head(toronto_shelters_clean)
```

```{r}
#| eval: false
#| echo: true

toronto_shelters_clean <-
  clean_names(toronto_shelters) |>
  mutate(occupancy_date = ymd(occupancy_date)) |> 
  select(occupancy_date, occupied_beds)

head(toronto_shelters_clean)
```

- p. 41: Remove a stray ":::".
- p. 66: "New Project$dots" should be "New Project...".
- p. 138: `scale_color_brewer(palette = "Set1")` is unnecesary and should be removed.
- p. 138: The figure caption should refer to inflation not unemployment.
- p. 154: Q9 is missing a "work" after the code chunk and before the "if".
- p. 347: The "Exploratory data analysis" chapter of *R for Data Science* is 11, not 12.
- p. 353: Fix a "the the".
- p. 355: "...results in an estimate of 5,814, both of which are too low." should be "...results in an estimate of 11,197, the former is too low and the latter is too high."
- p. 371: The sentence that references Figure 11.11a was confusing and needs to more clearly reference the figure. 
- p. 587: The link should be: https://fivethirtyeight.com/features/police-misconduct-costs-cities-millions-every-year-but-thats-where-the-accountability-ends/

## Updates

While I have tried to limit changes to what was printed, the following have been made:

**2023-07-24**

- Updated links and citations for: @barrett2021; @democratizingr; @chase2020; @kasymatching; and @matmortality.

**2023-08-10**

- Added a reference to @beautifulcode.

**2023-09-18**

- Updated links and citations for: @barrett2021.

**2023-11-12**

- Updated links and citations for: @barrett2021, again.
- Alberta updated their COD file so that it now needs a skip. 

**2023-01-10**

- Mentioned Python in the Introduction.
- Adding Prediction as a chapter.
- Updated from `get_stamenmap()` to `get_stadiamap()`.
- Added an "Installation and set-up" appendix chapter.