---
engine: knitr
---

# Errata and updates {#sec-errata}

I am grateful for the corrections and suggestions of: 
Crystal Lewis, and 
Zak Varty.

## Errata

The following errata exist in the print version, but have been updated in the online version. If you notice an error not mentioned below, please submit an [issue](https://github.com/RohanAlexander/telling_stories/issues) or send an email: rohan.alexander@utoronto.ca.

- p. 20: Added "packages" to "use the `tidyverse` and `janitor` packages."
- p. 34: `"daily-shelter-overnight-service-occupancy-capacity-2021"` should be `"daily-shelter-overnight-service-occupancy-capacity-2021.csv"` (note the ".csv" added).
- p. 34: Replace the first code chunk with the second:

```{r}
#| eval: false
#| echo: true

toronto_shelters_clean <-
     clean_names(toronto_shelters) |>
     select(occupancy_date, id, occupied_beds)

head(toronto_shelters_clean)
```

```{r}
#| eval: false
#| echo: true

toronto_shelters_clean <-
  clean_names(toronto_shelters) |>
  mutate(occupancy_date = ymd(occupancy_date)) |> 
  select(occupancy_date, occupied_beds)

head(toronto_shelters_clean)
```

- p. 41: Removed a stray ":::".
- p. 66: "New Project$dots" should be "New Project...".
- p. 587: The link should be: https://fivethirtyeight.com/features/police-misconduct-costs-cities-millions-every-year-but-thats-where-the-accountability-ends/

## Updates

While I have tried to limit changes to what was printed, the following have been made:

### 2023-07-24

- Updated links and citations for: @barrett2021; @democratizingr; @chase2020; @kasymatching; and @matmortality.



